Learning Reusable Components for Neurally--Guided Bayesian Program Learning

### 2018-05-12 nips_2018.tex
intro

\begin{abstract}
  Successful approaches to program induction require a hand-engineered
  domain-specific language (DSL), constraining the space of allowed
  programs and imparting prior knowledge of the domain.  We contribute
  a program induction algorithm called \system that learns a DSL and
  trains a neural network to efficiently search for programs in the
  learned DSL.  We use our model to build circuits, solve
  symbolic regression problems, edit strings, and synthesize functions on lists, in each case
  showing the model learns a domain-specific vocabulary for expressing
  solutions to problems in the domain.
\end{abstract}

@@

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
inventory of restricted programming primitives, encoding domain-specific
knowledge about the space of programs. In
practice
we often have only a few input/output examples for each
program to be induced, and thus success often hinges on having a good
DSL that provides a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered domain-specific languages?

We propose \emph{learning} the DSL. We consider the setting where we
have a collection of related programming tasks, each specified by a
set of input/output examples. We do not assume that the tasks are
annotated with ground-truth programs. We typically will not try to
learn a DSL completely from scratch, but rather to start from a weaker
or more general set of primitives and
construct a richer, more powerful, and better-tuned DSL.

Our algorithm is called \system because it is based on a
novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
between ``wake'' and ``sleep'' phases to achieve three goals: finding
programs that solve tasks; creating a DSL by discovering
and reusing domain-specific subroutines; and training a neural network
that efficiently guides search for programs in the DSL. The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
network ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

@@

Viewed as a probabilistic inference problem,
our goal is to jointly infer a single DSL (written $\mathcal{D}$)
and many programs (written $p$), one for each set of inputs/outputs (\emph{tasks}, written $x$).
A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a distribution over programs.
Fig.~\ref{graphicalModel}(a) diagrams this inference problem as a hierarchical Bayesian model.
Inference in this model is
difficult because the programs are unobserved,
and so we must solve a hard search problem to recover them. To make
search tractable we learn a bottom-up \emph{recognition
  model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
The recognition model $q(\cdot )$ is a neural
network that regresses from input/output pairs to a distribution over
programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
inference
scheme~\cite{le2016inference}.
 The neural recognition model and the
generative model embodied in the DSL jointly train each other, as they
iteratively learn to solve more programming tasks.

\begin{figure}
  \centering

  \begin{tikzpicture}

  \node[latent] at (0.5,3) (d){$\mathcal{D}$};
  \node[latent] at (1.5,3) (t){$\theta$};
  \node[latent] at (1,1.75) (z){$p$};
%  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
  \node[obs] at (1,0.5) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(z)(x)}{$x\in X$};
  \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative model};

  \begin{scope}[shift = {(0.7,0)}]  
  \node[obs] at (3.5,3) (dx){$\mathcal{D}$};
  \node[latent] at (3.5,1.9) (zp){$p$};
  \node[latent] at (5,1.9) (tx){$\theta^{(x)}$};
  \node[obs] at (3.5,0.5) (xp) {$x$};
  \plate {}{(zp)(xp)(tx)}{$x\in X$};
  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
  \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
  \draw [->,red] (tx.west) -- (zp.east);
  \draw [->,red] (dx.south) -- (zp.north);

  \node at (nn) {
    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}
  };
  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };

  \end{scope}

  \end{tikzpicture}
  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$. From each program we observe a task $x$ (program inputs/outputs). (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
    $\theta^{(x)}$. Black/red arrows correspond to the generative/recognition models.}\label{graphicalModel}
\end{figure}

@@

We apply \system to four domains:
 building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
 For each of these we deliberately provide an impoverished
 set of programming primitives,
 and show that our algorithm discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
 \begin{table}[b]
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   Circuits&\code{(nand (nand a a) (nand b b))}\\
   &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{table}

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration from the Wake/Sleep
algorithm~\cite{hinton1995wake}, as well as the
Exploration-Compression algorithm (EC) for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Wake/Sleep
alternates between training a generative model on samples from a
recognition model (our $q$), and training a recognition model on
samples from the generative model (our $\mathcal{D},\theta$).  EC
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We combine these ideas into an inference strategy that iterates
through three steps: a \textbf{Wake} cycle uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Sleep-G} and \textbf{Sleep-R} cycles update the DSL and the
recognition model, respectively.  Crucially, these steps bootstrap 
off each other (Fig.~\ref{feeding}):

@@

\begin{figure}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};

    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
\end{figure}

@@

\subsection{Probabilistic Framing}\label{mathematicalFraming}

@@

Recall from Sec.~\ref{mathematicalFraming} that we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$.
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{align}
  \nonumber
    \log \probability[\mathcal{D}] + \argmax_{\theta}&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\\
  &+ \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{align}

We induce a DSL by searching locally through the space of DSLs,
proposing small local changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting subexpressions from
programs already in the frontier.
These extracted subexpressions
are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
which then become new functions in the DSL.
The idea of storing and reusing
fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.

We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
\begin{align}
  \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
  \probability[\theta|\mathcal{D}]& = \text{Dir}(\theta|\alpha)
\end{align}
where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
$\lambda$ is a hyperparameter that acts as a regularizer on the size of the DSL,
and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

@@

\subsection{Boolean Circuits}
We consider the problem of learning
to build  Boolean circuits out of logic gates.
If we gave our agent the full repertoire of
circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
Instead, we gave \system only a NAND gate,
and tasked it with building a set of 250 random circuits made out of
AND, NOT, and OR gates, along with a multiplexer.
\system observes the input/output truth table of the circuits,
and we define the likelihood 
$  \probability[x|p]\triangleq\indicator\left[p\text{ has  the same truth table as }x \right]$,
where $x$ is a truth table and $p$ a program implementing a circuit.

This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
%% For this domain our recognition model is a simple multilayer perceptron 
%% whose inputs are the input/output examples.
\system learns to build these
circuits by creating new subroutines that correspond to
logic gates and multiplexers (Fig.~\ref{multiplexer}).
We will show that the neural recognition model
assists in building these circuits (Sec.~\ref{quantitative}),
suggesting that an agent can glean something about the structure of a circuit just by looking at its truth table.

\begin{figure}\centering
  \begin{circuitikz} \draw
    (0,0) node (x) {$x$}
    (x)++(0,-0.3) node (y) {$y$}
    (y)++(0,-0.5) node (z) {$z$}

    (x)++(1.7,-0.14) node[nand port, scale = 0.5] (xy) {}
    (xy) ++ (0,-1) node[nand port, scale = 0.5] (xz) {}
    (xz) ++ (1,0.2) node[nand port, scale = 0.5] (zxz) {}
    (zxz) ++ (1,0.2) node[nand port, scale = 0.5] (m) {}
    (x) -- (xy.in 1)
    (y) -- (xy.in 2)

    (x) -| (xz.in 1)
    (z.east) -- (xz.in 2)

    (xz.out) -- (zxz.in 2)
    (z.east) |- (zxz.in 1)

    (zxz.out) -- (m.in 2)
    (xy.out) |- (m.in 1)

    (m)++(1.2,0) node (l) {\textsc{MUX}$(x,y,z)$}
    ;
%% (0,0) node[and port] (myand2) {}
%% (2,1) node[xnor port] (myxnor) {}
%% (myand1.out) -- (myxnor.in 1)
%% (myand2.out) -- (myxnor.in 2);
  \end{circuitikz}
  \begin{circuitikz} \draw
    (0,0) node (a) {$a$}
    (a)++(0,-0.7) node (b) {$b$}

    (a)++(1,-0) node[nand port, scale = 0.5] (ab1) {}
    (b)++(1,-0) node[nand port, scale = 0.5] (ab2) {}
    (ab1)++(1,-0.35) node[nand port, scale = 0.5] (o) {}

    (a.east) -| (ab1.in 1)
    (a.east) -| (ab1.in 2)
    (b.east) -| (ab2.in 1)
    (b.east) -| (ab2.in 2)

    (ab1.out) -- (o.in 1)
    (ab2.out) -- (o.in 2)

    (o)++(0.8,0) node (l) {\textsc{OR}$(a,b)$}
    ;
  \end{circuitikz}
  \begin{circuitikz}     \node(m)[draw] at (1,-1) {MUX}; \draw
    (0,0) node (w) {$w$}
    (w)++(0,-0.5) node (x) {$x$}
    (x)++(0,-0.5) node (y) {$y$}
    (y)++(0,-0.5) node (z) {$z$}

    (z.east) -| (m.south)
    (y.east) -| (m.west)
    (x.east) -| (m.north)

    (y)++(2.5,0.13) node[or port, scale = 0.5] (z) {}
    (m.east) -- (z.in 2)
    (w.east) -| (z.in 1)
    ;
  \end{circuitikz}
\caption{\system discovers multiplexers (above) and OR gates (bottom left), which it uses to make more complicated circuits (bottom right) that it was unable to build without discovering these new primitives.}\label{multiplexer}
  \end{figure}

%% \begin{figure}
%%   \includegraphics[width = \columnwidth]{figures/circuit.png}
%%   \caption{\system learning to build Boolean circuits. We start out with only NAND gates and learn DSL that contains ...}\label{circuitLearningCurve}
%% \end{figure}

>>>>

\begin{abstract}
  Successful approaches to program induction require a hand-engineered
  domain-specific language (DSL), constraining the space of allowed
  programs and imparting prior knowledge of the domain.  We contribute
  a program induction algorithm called \system that learns a DSL while
  jointly training a neural network to efficiently search for programs
  in the learned DSL.  We use our model to solve symbolic regression
  problems, edit strings, and synthesize functions on lists, showing
  how the model learns a domain-specific library of program components
  for expressing solutions to problems in the domain.
\end{abstract}

@@

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain-Specific Language} (DSL). DSLs contain an
inventory of restricted programming primitives, encoding domain-specific
knowledge about the space of programs. In
practice
we often have only a few input/output examples for each
program to be induced, and thus success often hinges on having a good
DSL that provides a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered domain-specific languages?

We propose \emph{learning} the DSL by inducing a library of domain--specific
subroutines. We consider the setting where we have a collection of
related programming tasks, each specified by a set of input/output
examples. %% We do not assume that the tasks are annotated with
%% ground-truth programs.
Start from a weaker or more
general library of primitives, we give an algorithm for constructing a richer, more powerful,
and better-tuned DSL.
Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding)
because it iterates between three different phases: an
\textbf{Explore} phase uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} phase
modifies the DSL by discovering regularities in the programs found by
the previous Explore phase; and a \textbf{Compile} phase, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
network ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

Because any model may be encoded as a (deterministic or probabilistic) program,
we carefully delineate the scope of program learning problems considered here.
We think of \system as learning to solve the kinds of problems that humans
can solve relatively quickly -- once they acquire the relevant domain expertise.
These correspond to short programs -- once you have the right DSL.
Even with the right DSL, program search may be intractable, so
we amortize the cost of program search by training a neural network to
assist the search procedure.

@@

%% \begin{figure}
%%   \centering

%%   \begin{tikzpicture}

%%   \node[latent] at (0.5,3) (d){$\mathcal{D}$};
%%   \node[latent] at (1.5,3) (t){$\theta$};
%%   \node[latent] at (1,1.75) (z){$p$};
%% %  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
%%   \node[obs] at (1,0.5) (x) {$x$};
%%   \edge {z}{x};
%%   \edge {d,t}{z};
%%   \plate {}{(z)(x)}{$x\in X$};
%%   \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative model};

%%   \begin{scope}[shift = {(0.7,0)}]  
%%   \node[obs] at (3.5,3) (dx){$\mathcal{D}$};
%%   \node[latent] at (3.5,1.9) (zp){$p$};
%%   \node[latent] at (5,1.9) (tx){$\theta^{(x)}$};
%%   \node[obs] at (3.5,0.5) (xp) {$x$};
%%   \plate {}{(zp)(xp)(tx)}{$x\in X$};
%%   \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
%%   \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
%%   \draw [->,red] (tx.west) -- (zp.east);
%%   \draw [->,red] (dx.south) -- (zp.north);

%%   \node at (nn) {
%%     \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
%%       \node[rectangle] at (1,1) {};
%%       \foreach \name / \y in {1,...,4}
%%           \node[neuron] (I-\name) at (0,-\y) {};
%%       \foreach \name / \y in {1,...,3}
%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
%%       \foreach \name / \y in {1,...,4}
%%           \node[neuron] (O-\name) at (2,-\y) {};
%%       \foreach \source in {1,...,4}
%%           \foreach \dest in {1,...,3}
%%               \draw [-latex] (I-\source) -- (H-\dest);
%%       \foreach \source in {1,...,3}
%%           \foreach \dest in {1,...,4}
%%               \draw [-latex] (H-\source) -- (O-\dest);
%%     \end{tikzpicture}
%%   };
%%   \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };

%%   \end{scope}

%%   \end{tikzpicture}
%%   \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$. From each program we observe a task $x$ (program inputs/outputs). (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
%%     $\theta^{(x)}$. Black/red arrows correspond to the generative/recognition models.}\label{graphicalModel}
%% \end{figure}

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we provide a generic
 set of programming primitives in a Lisp-like language,
 including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).

 \begin{wrapfig}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
   &\hspace{0.5cm} \emph{(add k to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{wrapfig}

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration primarily from the
Exploration-Compression algorithm for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We extend these ideas into an inference strategy that iterates
through three steps: an \textbf{Explore} cycle uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Compress} and \textbf{Compile} cycles update the DSL and the
recognition model, respectively.  Crucially, these steps bootstrap 
off each other (Fig.~\ref{feeding}):

@@

\begin{wrapfig}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compression)}} (DSL.30);
    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.150);
    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.west);

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a neural network called the \emph{recognition model}. Each of these steps bootstrap off of the others in an iterative inference algorithm.}  \label{feeding}
\end{wrapfigure}

@@

\subsection{Probabilistic Framing}\label{mathematicalFraming}
Viewed as a probabilistic inference problem,
our goal is to jointly infer a single DSL (written $\mathcal{D}$)
and many programs (written $p$), one for each set of inputs/outputs (\emph{tasks}, written $x$).
A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a distribution over programs.
Fig.~\ref{graphicalModel}(a) diagrams this inference problem as a hierarchical Bayesian model.
Inference in this model is
difficult because the programs are unobserved,
and so we must solve a hard search problem to recover them. To make
search tractable we learn a bottom-up \emph{recognition
  model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
The recognition model $q(\cdot )$ is a neural
network that regresses from input/output pairs to a distribution over
programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
inference
scheme~\cite{le2016inference}.
 The neural recognition model and the
generative model embodied in the DSL jointly train each other, as they
iteratively learn to solve more programming tasks.

@@

Recall from Sec.~\ref{mathematicalFraming} that we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$.
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small local changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting subexpressions from
programs already in the frontier.
These extracted subexpressions
are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
which then become new functions in the DSL.
The idea of storing and reusing
fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.

We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
\begin{equation}
    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\hspace{1cm}  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)
  \end{equation}

where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
$\lambda$ is a hyperparameter that acts as a regularizer on the size of the DSL,
and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

@@
